### 2021-11-23 paper/localpsf.tex
added minimum ellipsoid width functionality

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
}
\fi

@@

% REQUIRED
\begin{abstract}
	We present an efficient matrix-free method for approximating locally translation invariant operators that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in Schur complements methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems. The method computes impulse responses of the operator at a collection of scattered points, then interpolates these computed impulse responses to form a product-convolution approximation of the operator. The product-convolution approximation is converted to hierarchical matrix format in order to solve linear systems and perform other matrix operations efficiently. Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources. The key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. We apply the method to approximate Hessians in large-scale PDE-constrained inverse problems with highly informative data. Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation. Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
\end{abstract}

>>>>

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of locally translation invariant operators that have locally supported non-negative integral kernels, with application to Hessians in PDE-constrained inverse problems},
  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
}
\fi

@@

% REQUIRED
\begin{abstract}
	We present a fast matrix-free method for computing hierarchical matrix (H-matrix) approximations of locally translation invariant operators that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems. Converting an operator to H-matrix format is highly desirable, because fast scalable H-matrix methods can be used to add, multiply, invert, and factorize H-matrices, perform matrix-vector products, solve linear systems, and perform other useful matrix operations. The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$. Classical H-matrix construction methods require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible in large-scale problems when the operator is only accessible via matrix-vector products (``matrix-free''). We overcome this problem by computing impulse responses of the operator and its transpose at a collection of scattered points, then interpolating shifted versions of these impulse responses on-the-fly to evaluate the integral kernel rapidly. Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources. The key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. We apply the method to approximate Hessians in large-scale PDE-constrained inverse problems with highly informative data. Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation. We are able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
\end{abstract}
